---
title: "markdowntest"
author: "Kelly Okech"
date: "Monday, June 08, 2015"
output: html_document
---
This is a report detailing the steps I took in answering assignment 1 in reproducible research class. 

So, to begin, we'll load the csv file and probably try to explore the data in it

```{r}
activity <- read.table("activity.csv", header=T, sep=",")

# convert date to date data type
activity$date <- as.Date(activity$date)
str(activity)

```

Doing away with missing values

```{r}
# Doing away with na values
activity1 <- na.omit(activity)

#Checking new structure
str(activity1)
```

Question 1: total mean number of steps taken per day

Grouping the rows into 24hour intervals and determining their sum
```{r}
daily.steps <- rowsum(activity1$steps, format(activity1$date, '%Y-%m-%d')) 
daily.steps <- data.frame(daily.steps) 
names(daily.steps) <- ("steps") 
```

Plot a histogram of steps variable to explore the data further
```{r}
hist(daily.steps$steps, main = "daily steps", xlab = "Total daily steps")
abline(v = mean(daily.steps$steps), col = "Blue")
```

Calculate the mean value
```{r}
mean(daily.steps$steps)
```

Calculate the median value
```{r}
median(daily.steps$steps)
```

Question 2: Find the average daily activity pattern

```{r}
library(plyr)
# average steps for each of 5-minute interval during a 24-hour period
interval <- ddply(activity1,~interval, summarise, mean=mean(steps))
```

Time series plot of the intervals and average steps

```{r}
library(ggplot2)
qplot(x=interval, y=mean, data = interval,  geom = "line",
      xlab="5-Minute Interval",
      ylab="Step Count",
      main="Average Number of Steps in 24 hours across all the days"
      )
```

Finding out which 5 minute intervals has the highest number of steps

```{r}
interval[which.max(interval$mean), ]
```

Question 3: imputing missing values

Finding th total number fo missing values in the data set.

```{r}
#use the str function on data set activity1.Total number of na values is indicated somewhere in there as 2304
str(activity1)
```

alternatively, you can run this code to find out how many missing values are in the variable 'steps'

```{r}
sum(is.na(activity$steps))
```

We need to fil in the missing data

```{r}
fill <- activity


#check for missing value
sum(is.na(fill))
```

Make another histogram to compare with the first one

```{r}
newsteps<-tapply(fill$steps, fill$date, sum)
```






